infer: An R package for tidyverse-friendly statistical inference

### 2021-07-11 figs/paper/paper.tex
note Downey's "there is only one test"

\texttt{infer} is an R package for randomization-based hypothesis
testing, naturalizing an intuitive understanding of statistical
inference via a unified and expressive grammar. Four functions provide
functionality encompassing a large swath of basic frequentist
statistical inference, abstracting away details of specific tests and
shifting the focus of the analyst to the observed data and the processes
that generated it. Such a grammar lends itself to applications in
teaching, data pedagogy research, applied scientific research, and
advanced predictive modeling. For one, the principled approach of the
package has made it an especially good fit for teaching introductory
statistics and data science (Baumer, Bray, Çetinkaya-Rundel, \& Hardin,
2020; Çetinkaya-Rundel \& Ellison, 2021; Ismay \& Kim, 2019) and
research in data pedagogy (Fergusson \& Pfannkuch, 2021; Loy, 2021).
Further, the package has already seen usage in a number of published
scientific applications (Ask et al., 2021; Fallon \& Hinds, 2021; McLean
et al., 2021). Finally, the package integrates with the greater
tidymodels collection of packages, a burgeoning software ecosystem for
tidyverse-aligned predictive modeling used across many modern research
and industrial applications (Kuhn \& Wickham, 2020). To date, the
package has been downloaded more than 400,000 times.

@@

  \begin{itemize}
  \tightlist
  \item
    \texttt{shade\_p\_value()} situates the observed statistic in the
    null distribution, shading the region as or more extreme.
  \end{itemize}
\item
  \texttt{get\_p\_value()} calculates a p-value via the juxtaposition of
  the test statistic and null distribution.
\end{itemize}

@@

Beyond this, the package offers:

@@

\texttt{broom} and \texttt{parameters} convert model objects to unified
output formats, though do not provide methods for fitting models,
describing null distributions, performing bootstrapping, or calculating
summary statistics from tabular data. \texttt{statsexpressions}, and
adjacent packages in the \texttt{easystats} ecosystem, implement
wrappers with consistent interfaces for theory-based hypothesis tests.
At a higher level, the structure of each of these packages is defined by
model types, where each model type has its own associated function
and/or model object. In contrast, \texttt{infer} is structured around
four functions, situating statistics and model types within a more
abstracted grammar. \texttt{infer}'s initial release on the
Comprehensive R Archive Network predated that of each of the packages
mentioned above, with the exception of \texttt{broom}, a package
situated in the same \texttt{tidymodels} ecosystem (R Core Team, 2021).

@@

>>>>

\texttt{infer} is an R package for randomization-based hypothesis
testing, naturalizing an intuitive understanding of statistical
inference via a unified and expressive grammar. Four functions provide
functionality encompassing a large swath of basic frequentist
statistical inference, abstracting away details of specific tests and
shifting the focus of the analyst to the observed data and the processes
that generated it. Such a grammar lends itself to applications in
teaching, data pedagogy research, applied scientific research, and
advanced predictive modeling. For one, the principled approach of the
\texttt{infer} package has made it an especially good fit for teaching
introductory statistics and data science (Baumer, Bray,
Çetinkaya-Rundel, \& Hardin, 2020; Çetinkaya-Rundel \& Ellison, 2021;
Ismay \& Kim, 2019) and research in data pedagogy (Fergusson \&
Pfannkuch, 2021; Loy, 2021). Further, the package has already seen usage
in a number of published scientific applications (Ask et al., 2021;
Fallon \& Hinds, 2021; McLean et al., 2021). Finally, the package
integrates with the greater tidymodels collection of packages, a
burgeoning software ecosystem for tidyverse-aligned predictive modeling
used across many modern research and industrial applications (Kuhn \&
Wickham, 2020). To date, the package has been downloaded more than
400,000 times.

@@

  \begin{itemize}
  \tightlist
  \item
    \texttt{shade\_p\_value()} situates the observed statistic in the
    null distribution, shading the region as or more extreme.
  \end{itemize}
\item
  \texttt{get\_p\_value()} calculates a p-value via the juxtaposition of
  the test statistic and the null distribution.
\end{itemize}

@@

Beyond this, the \texttt{infer} package offers:

@@

\texttt{broom} and \texttt{parameters} convert model objects to unified
output formats, though they do not provide methods for fitting models,
describing null distributions, performing bootstrapping, or calculating
summary statistics from tabular data. \texttt{statsexpressions}, and
adjacent packages in the \texttt{easystats} ecosystem, implement
wrappers with consistent interfaces for theory-based hypothesis tests.
At a higher level, the structure of each of these packages is defined by
model types, where each model type has its own associated function
and/or model object. In contrast, \texttt{infer} is structured around
four functions, situating statistics and model types within a more
abstracted grammar.\footnote{This grammar follows from Allen Downey's
  ``there is only one test'' framework (Downey, 2011).} \texttt{infer}'s
initial release on the Comprehensive R Archive Network predated that of
each of the packages mentioned above, with the exception of
\texttt{broom}, a package situated in the same \texttt{tidymodels}
ecosystem (R Core Team, 2021).

@@

\leavevmode\hypertarget{ref-downey2011think}{}%
Downey, A. B. (2011). \emph{Think stats}. " O'Reilly Media, Inc.".
